# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
* Step 1: Read the data from the .zip file into our environment
```{r, echo=TRUE}
pa1.data <- read.csv(unzip("activity.zip"), header=TRUE,
                     colClasses = c("integer","Date","integer"));
```

## What is mean total number of steps taken per day?

* Step 1 : Generate Histogram of number of steps taken by Date
```{r, echo=TRUE}
hist(sapply(split(pa1.data$steps, pa1.data$date), sum), 
     xlab="Steps", main="Steps Taken by Date");
```

* Step 2: Generate data frame with mean and median by date, then use kable to format it as a table

```{r, echo=TRUE, results="asis"}
pa1.meanByDate <- aggregate(steps ~ date, data=pa1.data, 
                            FUN=function(x) c(Mean=mean(x), Median=median(x)));
colnames(pa1.meanByDate) <- c("Date", "Mean Steps", "Median Steps");
kable(data.frame(Date=pa1.meanByDate$date, 
                 Mean=pa1.meanByDate$steps[, 1], 
                 Median=pa1.meanByDate$steps[, 2]));
```

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
